---
output:
  pdf_document:
    citation_package: natbib
    keep_tex: no
    latex_engine: pdflatex
fontfamily: mathpazo
fontsize: 11pt
geometry: margin=1in
header-includes: 
- \linespread{1.15}
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
memorandum: yes
graphics: yes
width: 0.3
---

```{r, include = FALSE, message = FALSE}
library(tidyverse)
library(dplyr)
library(psych)
library(ggrepel)
library(ggcorrplot)
library(reshape2)
library(knitr)
library(rempsyc)
library(kableExtra)
library(GGally)
library(formattable)
library(memisc)
library(stargazer)
library(haven)
library(lmtest)
library(sandwich)
```

```{r, include = FALSE}
econ_free <- read_csv("EconomicFreedomCountries.csv", show_col_types = FALSE)
```

```{r, include = FALSE}
glimpse(econ_free)
```

```{r, include = FALSE}
econ_free_clean <- dplyr::rename(econ_free, iso_code = ISO_code,
                  country = countries,
                  econ_freedom_score = 'ECONOMIC FREEDOM',
                  gov_consumption = `1a_government_consumption`,
                  transfers = `1b_transfers`,
                  gov_enterprises = `1c_gov_enterprises`,
                  top_marg_tax_rate = `1d_top_marg_tax_rate`,
                  gov_size = `1_size_government`,
                  judicial_indep = `2a_judicial_independence`,
                  impartial_courts = `2b_impartial_courts`,
                  protec_prop_rights = `2c_protection_property_rights`,
                  military_interf = `2d_military_interference`,
                  integrity_legal_syst = `2e_integrity_legal_system`, 
                  legal_enforc_contracts = `2f_legal_enforcement_contracts`,
                  restr_sale_real_prop = `2g_restrictions_sale_real_property`,
                  police_reliability = `2h_reliability_police`,
                  busin_costs_crime = `2i_business_costs_crime`,
                  gender_adjust = `2j_gender_adjustment`,
                  property_rights = `2_property_rights`,
                  money_growth = `3a_money_growth`,
                  std_inflation = `3b_std_inflation`,
                  inflation = `3c_inflation`,
                  foreign_currency = `3d_freedom_own_foreign_currency`,
                  sound_money = `3_sound_money`,
                  tariffs = `4a_tariffs`,
                  regu_trade_barriers = `4b_regulatory_trade_barriers`,
                  black_market = `4c_black_market`,
                  control_movement = `4d_control_movement_capital_ppl`,
                  trade = `4_trade`,
                  credit_market_reg = `5a_credit_market_reg`,
                  labor_market_reg = `5b_labor_market_reg`,
                  business_reg = `5c_business_reg`,
                  regulation = `5_regulation`)
```

```{r, include = FALSE}
econ_free_vars_q1 <- econ_free_clean %>%
  dplyr::select(rank, gov_size, property_rights, sound_money, trade, regulation)

econ_free_vars_q2 <- econ_free_clean %>%
  dplyr::select(top_marg_tax_rate, inflation, money_growth, std_inflation, regu_trade_barriers, business_reg)
```

\begin{center}\textbf{MEMORANDUM}\end{center}

\textbf{FROM}: Patrick Cavanaugh and Eric Elizes  
\textbf{TO}: Mx. Gourinchas, Chief Economist of the IMF  
\textbf{SUBJECT}: Results of Economic Freedom analysis  
\textbf{DATE}: January 28, 2023

\par\noindent\rule{\textwidth}{0.4pt}

## SUMMARY 

The application of many regression methods found that (1) government size has a large negative impact on rank, but not as large as property rights and sound money, and (2) trade and regulation have large positive impacts on rank. These findings are best illustrated by the multivariate regression with interactions and the log-log model.

```{r, echo=FALSE}
simple_reg <- lm(formula = rank ~ gov_size, data = econ_free_clean)
simple_reg <- coeftest(simple_reg, vcov. = vcovHC(simple_reg, type = 'HC1'))

multivar_reg <- lm(formula = rank ~ gov_size + property_rights + sound_money + trade + regulation, data = econ_free_clean)
multivar_reg <- coeftest(multivar_reg, vcov. = vcovHC(multivar_reg, type = 'HC1'))

fixed_eff_est <- lm(formula = rank ~ gov_size + property_rights + sound_money + trade + regulation + country , data = econ_free_clean)
fixed_eff_est <- coeftest(fixed_eff_est, vcov. = vcovHC(fixed_eff_est, type = 'HC1'))

int_term <- lm(formula = rank ~ gov_size + property_rights + sound_money + trade + regulation + country  + trade * regulation , data = econ_free_clean)
int_term <- coeftest(int_term, vcov. = vcovHC(int_term, type = 'HC1'))

multi_1 <- lm(formula = rank ~ gov_size + property_rights + trade + regulation + country , data = econ_free_clean)
multi_1 <- coeftest(multi_1, vcov. = vcovHC(multi_1, type = 'HC1'))

multi_2 <- lm(formula = rank ~ gov_size + property_rights + sound_money + regulation + country , data = econ_free_clean)
multi_2 <- coeftest(multi_2, vcov. = vcovHC(multi_2, type = 'HC1'))

log_lin <- lm(formula = log(rank) ~ gov_size + property_rights + sound_money + trade + regulation + country , data = econ_free_clean)
log_lin <- coeftest(log_lin, vcov. = vcovHC(log_lin, type = 'HC1'))

econ_free_clean2 <- econ_free_clean
econ_free_clean2["sound_money"][econ_free_clean2["sound_money"] == 0.000000] <- NA
econ_free_clean2["trade"][econ_free_clean2["trade"] == 0.000000] <- NA

log_log <- lm(formula = log(rank) ~ log(gov_size) + log(property_rights) + log(sound_money) + log(trade) + log(regulation) + country, data = econ_free_clean2, na.action = na.omit)
log_log <- coeftest(log_log, vcov. = vcovHC(log_log, type = 'HC1'))
```

The following legend includes the abbreviations for Table 2: Regression Results. The regressions will be more thoroughly explained throughout the paper.

```{r echo=FALSE, warning=FALSE}
my_table <- data.frame(Abbreviation = c("SR", "MR", "FE", "IT", "MA1", "MA2", "Log-Lin", "Log-Log"),
                       Full = c("Simple Regression", "Multivariate Regression", "Fixed Effects Estimator", "Interaction Term", "Multicollinearity Analysis 1", "Multicollinearity Analysis 2", "Log-Linear Model", "Log-Log Model"))

kable(my_table, caption = "Regressions Legend") %>%
  kable_classic(latex_options = "HOLD_position")
```