---
output:
  pdf_document:
    citation_package: natbib
    keep_tex: no
    latex_engine: pdflatex
  html_document:
    df_print: paged
fontfamily: mathpazo
fontsize: 11pt
geometry: margin=1in
header-includes: \linespread{1.15}
memorandum: yes
graphics: yes
width: 0.3
---

```{r, include = FALSE, message = FALSE}
library(tidyverse)
library(dplyr)
library(psych)
library(ggrepel)
library(ggcorrplot)
library(reshape2)
library(knitr)
library(rempsyc)
library(kableExtra)
library(GGally)
library(formattable)
library(stargazer)
```

```{r, include = FALSE}
econ_free <- read_csv("EconomicFreedomCountries.csv", show_col_types = FALSE)
```

```{r, include = FALSE}
glimpse(econ_free)
```

```{r, include = FALSE}
econ_free_clean <- rename(econ_free, iso_code = ISO_code,
                  country = countries,
                  econ_freedom_score = 'ECONOMIC FREEDOM',
                  gov_consumption = `1a_government_consumption`,
                  transfers = `1b_transfers`,
                  gov_enterprises = `1c_gov_enterprises`,
                  top_marg_tax_rate = `1d_top_marg_tax_rate`,
                  gov_size = `1_size_government`,
                  judicial_indep = `2a_judicial_independence`,
                  impartial_courts = `2b_impartial_courts`,
                  protec_prop_rights = `2c_protection_property_rights`,
                  military_interf = `2d_military_interference`,
                  integrity_legal_syst = `2e_integrity_legal_system`, 
                  legal_enforc_contracts = `2f_legal_enforcement_contracts`,
                  restr_sale_real_prop = `2g_restrictions_sale_real_property`,
                  police_reliability = `2h_reliability_police`,
                  busin_costs_crime = `2i_business_costs_crime`,
                  gender_adjust = `2j_gender_adjustment`,
                  property_rights = `2_property_rights`,
                  money_growth = `3a_money_growth`,
                  std_inflation = `3b_std_inflation`,
                  inflation = `3c_inflation`,
                  foreign_currency = `3d_freedom_own_foreign_currency`,
                  sound_money = `3_sound_money`,
                  tariffs = `4a_tariffs`,
                  regu_trade_barriers = `4b_regulatory_trade_barriers`,
                  black_market = `4c_black_market`,
                  control_movement = `4d_control_movement_capital_ppl`,
                  trade = `4_trade`,
                  credit_market_reg = `5a_credit_market_reg`,
                  labor_market_reg = `5b_labor_market_reg`,
                  business_reg = `5c_business_reg`,
                  regulation = `5_regulation`)
```

```{r, include = FALSE}
econ_free_vars_q1 <- econ_free_clean %>%
  select(rank, gov_size, property_rights, sound_money, trade, regulation)

econ_free_vars_q2 <- econ_free_clean %>%
  select(top_marg_tax_rate, inflation, money_growth, std_inflation, regu_trade_barriers, business_reg)
```

\begin{center}\textbf{MEMORANDUM}\end{center}

\textbf{FROM}: Patrick Cavanaugh and Eric Elizes  
\textbf{TO}: Mx. Gourinchas, Chief Economist of the IMF  
\textbf{SUBJECT}: Results of Economic Freedom analysis  
\textbf{DATE}: January 28, 2023

\par\noindent\rule{\textwidth}{0.4pt}

## SUMMARY 

Summarize basic findings. The regression model that is most appropriate for 

These results are summarized in the following table:

```{r mylatextable, results = "asis", warning = FALSE, echo=FALSE}
simple_reg <- lm(formula = rank ~ gov_size, data = econ_free_clean)
multivar_reg <- lm(formula = rank ~ gov_size + property_rights + sound_money + trade + regulation, data = econ_free_clean)
fixed_eff_est <- lm(formula = rank ~ gov_size + property_rights + sound_money + trade + regulation + country , data = econ_free_clean)
int_term <- lm(formula = rank ~ gov_size + property_rights + sound_money + trade + regulation + country  + trade * regulation , data = econ_free_clean)
multi_1 <- lm(formula = rank ~ gov_size + property_rights + trade + regulation + country , data = econ_free_clean)
multi_2 <- lm(formula = rank ~ gov_size + property_rights + sound_money + regulation + country , data = econ_free_clean)
log_lin <- lm(formula = log(rank) ~ gov_size + property_rights + sound_money + trade + regulation + country , data = econ_free_clean)

econ_free_clean2 <- econ_free_clean
econ_free_clean2["sound_money"][econ_free_clean2["sound_money"] == 0.000000] <- NA
econ_free_clean2["trade"][econ_free_clean2["trade"] == 0.000000] <- NA

log_log <- lm(formula = log(rank) ~ log(gov_size) + log(property_rights) + log(sound_money) + log(trade) + log(regulation) + country, data = econ_free_clean2, na.action = na.omit)

stargazer(simple_reg, multivar_reg, fixed_eff_est, int_term, multi_1, multi_2, log_lin, log_log, 
          omit = 'country', 
          type = 'text',
          title = "What contributes to the rank of a country?",
          align = TRUE,
          column.labels = c("Simple Regression", "Multivariate Regression", " Fixed Effects Estimator", "Multivariate Regression with Interaction Term", "Multicollinearity Analysis 1", "Multicollinearity Analysis 2", "Log-Linear Model", "Log-Log Model"),
          omit.stat = c("ser", "adj.rsq", "f"))
```

## SIMPLE REGRESSION

When looking at what impacts the rank of countries, a variable of interest would be government size. As the size of the government increases, its is predicted that the rank would fall, indicating a lower economic freedom score. This is because as the size of government increases, by way of state ownership of assets and top marginal tax rate, economic freedom is limited. The simple regression that was run is as follows: 

$$ \hat{rank} = b_0 + b_1 * government \ size $$

The results of this regression are included below:
```{r, echo=FALSE}
summary(simple_reg)
```

From the results of the regression, a 1 unit increase in government size is associated with a 1.97 decrease in rank on average. The results of the simple regression agree with the prediction that as government size increases, rank decreases. This result was found to be statistically significant at the 0.001 significance level. This result is also economically significant as the size  of government can greatly affect the rank of countries.

## MULTIVARIATE REGRESSION

The multivariate regression that was run is as follows:

$$ \hat{rank} = b_0 + b_1 * government \ size + b_2 * property \ rights + b_3 * sound \ money + b_4 * trade + b_5 * regulation $$

The results of this regression are included below:

```{r, echo=FALSE}
summary(multivar_reg)
```

From the results of the regression, a 1 unit increase in government size is associated with a 4.59 decrease in rank on average, holding property rights, sound money, trade, and regulation constant. The results of the multivariate regression also agree with the prediction that as government size increases, rank decreases. This result was found to be statistically significant at the 0.001 significance level. Compared to the simple regression, this regression shows that government size has an even larger affect on rank. This result is even more economically significant as the size of government can greatly affect the rank of countries. The constant in this regression represents the rank value that would be predicted if all the independent variables were simultaneously equal to zero. This is not economically significant because it is not possible.

## FIXED EFFECTS ESTIMATOR

Being located in different countries The fixed effects estimator regression that was run is as follows:

$$ \hat{rank} = b_0 + b_1 * government \ size + b_2 * property \ rights + b_3 * sound \ money + b_4 * trade + b_5 * regulation + b_6 * country $$

The results of this regression are as follows:

```{r}
summary(fixed_eff_est)
```


## MULTIVARIATE REGRESSION WITH INTERACTION TERM

The multivariate regression with an interaction term that was run is as follows:

$$ \hat{rank} = b_0 + b_1 * government \ size + b_2 * property \ rights + b_3 * sound \ money + b_4 * trade + b_5 * regulation + b_6 * country + b_7 * trade * regulation $$

The results of this regression are as follows:

```{r}
summary(int_term)
```

## MULTICOLLINEARITY ANALYSIS

During the correlation analysis, it was noted that trade and sound money had the highest correlation between the explanatory variables with a correlation of 0.68. To further investigate the multicollinearity, the two following regressions were run:

$$ \hat{rank} = b_0 + b_1 * government \ size + b_2 * property \ rights + b_3 * trade + b_4 * regulation + b_5 * country $$

$$ \hat{rank} = b_0 + b_1 * government \ size + b_2 * property \ rights + b_3 * sound \ money + b_4 * regulation + b_5 * country $$

The results of this regression are as follows:

```{r, echo=FALSE}
summary(multi_1)

summary(multi_2)
```
 
 
 -3.7618 
 -3.86617
 -2.9223

## LOG-LINEAR MODEL

The log-linear regression that was run is as follows:

$$ log(\hat{rank}) = b_0 + b_1 * government \ size + b_2 * property \ rights + b_3 * sound \ money + b_4 * trade + b_5 * regulation + b_6 * country $$

The results of this regression are as follows:
```{r, echo=FALSE}
summary(log_lin) 
```

Based on the results of the regression, an in government size by one unit decreases rank by 8.67%. This was found to be statistically significant at the 0.001 significance level. This result is significantly significant, because a small change in government size contributes to a larger change on the change in rank.

### LOG-LOG MODEL

The log-log regression that was run is as follows:

$$ log(\hat{rank}) = b_0 + b_1 * log(government) \ size + b_2 * log(property \ rights) + b_3 * log(sound \ money) + b_4 * log(trade) + b_5 * log(regulation) + b_6 * country $$

The results of this regression are as follows:

```{r}
summary(log_log) 
```

Based on the results of the regression, a 1% increase in government size decreases the rank by 47.94%. This was found to be statistically significant at the 0.001 significance level. This result is significantly significant, because a small change in government size contributes to a large change on the change in rank.

