---
output:
  pdf_document:
    citation_package: natbib
    keep_tex: no
    latex_engine: pdflatex
fontfamily: mathpazo
fontsize: 11pt
geometry: margin=1in
header-includes: \linespread{1.15}
memorandum: yes
graphics: yes
width: 0.3
---

```{r, include = FALSE, message = FALSE}
library(tidyverse)
library(dplyr)
library(psych)
library(ggrepel)
library(ggcorrplot)
library(reshape2)
library(knitr)
library(rempsyc)
library(kableExtra)
library(GGally)
library(formattable)
```

```{r, include = FALSE}
econ_free <- read_csv("EconomicFreedomCountries.csv", show_col_types = FALSE)
```

```{r, include = FALSE}
glimpse(econ_free)
```

```{r, include = FALSE}
econ_free_clean <- rename(econ_free, iso_code = ISO_code,
                  country = countries,
                  econ_freedom_score = 'ECONOMIC FREEDOM',
                  gov_consumption = `1a_government_consumption`,
                  transfers = `1b_transfers`,
                  gov_enterprises = `1c_gov_enterprises`,
                  top_marg_tax_rate = `1d_top_marg_tax_rate`,
                  gov_size = `1_size_government`,
                  judicial_indep = `2a_judicial_independence`,
                  impartial_courts = `2b_impartial_courts`,
                  protec_prop_rights = `2c_protection_property_rights`,
                  military_interf = `2d_military_interference`,
                  integrity_legal_syst = `2e_integrity_legal_system`, 
                  legal_enforc_contracts = `2f_legal_enforcement_contracts`,
                  restr_sale_real_prop = `2g_restrictions_sale_real_property`,
                  police_reliability = `2h_reliability_police`,
                  busin_costs_crime = `2i_business_costs_crime`,
                  gender_adjust = `2j_gender_adjustment`,
                  property_rights = `2_property_rights`,
                  money_growth = `3a_money_growth`,
                  std_inflation = `3b_std_inflation`,
                  inflation = `3c_inflation`,
                  foreign_currency = `3d_freedom_own_foreign_currency`,
                  sound_money = `3_sound_money`,
                  tariffs = `4a_tariffs`,
                  regu_trade_barriers = `4b_regulatory_trade_barriers`,
                  black_market = `4c_black_market`,
                  control_movement = `4d_control_movement_capital_ppl`,
                  trade = `4_trade`,
                  credit_market_reg = `5a_credit_market_reg`,
                  labor_market_reg = `5b_labor_market_reg`,
                  business_reg = `5c_business_reg`,
                  regulation = `5_regulation`)
```

```{r, include = FALSE}
econ_free_vars_q1 <- econ_free_clean %>%
  select(rank, gov_size, property_rights, sound_money, trade, regulation)

econ_free_vars_q2 <- econ_free_clean %>%
  select(top_marg_tax_rate, inflation, money_growth, std_inflation, regu_trade_barriers, business_reg)
```

\begin{center}\textbf{MEMORANDUM}\end{center}

\textbf{FROM}: Patrick Cavanaugh and Eric Elizes  
\textbf{TO}: Mx. Gourinchas, Chief Economist of the IMF  
\textbf{SUBJECT}: Results of Economic Freedom analysis  
\textbf{DATE}: January 28, 2023

\par\noindent\rule{\textwidth}{0.4pt}

## SUMMARY 



## SIMPLE REGRESSION

```{r}

```



## MULTIVARIATE REGRESSION

```{r}

```



## FIXED EFFECTS ESTIMATOR

```{r}

```



## MULTIVARIATE REGRESSION WITH INTERACTION TERM

```{r}

```



## MULTICOLLINEARITY ANALYSIS



## LOG-LINEAR MODEL


### LOG-LOG MODEL


